\chapter{Future work}
\label{chap:future-work}

This work provides a solid base for future work on the capabilities and aspects of Web-Assembly in \gls{serverless} environment. This section will discuss some of the important areas that can be explored in the future works:
\begin{itemize}
    \item \textbf{Wasm GC}: The WebAssembly garbage collection proposal is still not finalized. Once the proposal is merged into the primary spec repository and memory-managed languages like Kotlin and C\# and Java adapt the WebAssembly garbage collection proposal, it would be interesting to evaluate the performance of these languages in serverless environment. The future work can also answer the question of whether memory-managed languages can be used in edge computing environment and how they perform in comparison to their non memory-managed counterparts like Rust and C++.
    \item \textbf{Web Frameworks}: The serverless area of WebAssembly can gain traction if popular web frameworks such as SvelteKit, Next.js, Nuxt.js, Solid, and many others could be hosted on WebAssembly-powered serverless platforms. We've observed that SvelteKit, along with possibly other web frameworks, is adopting the adapter pattern \cite{sveltecommunity_2023_adapter} to support multiple serverless platforms. Therefore, a future work can be to create an adapter for an existing WebAssembly platform, then assessing the feasibility and challenges associated with hosting web frameworks on these WebAssembly-powered serverless platforms.
    \item \textbf{Asynchronous with WASI}: Soon the WASI community will start with preview3. The preview3 \cite{gohman_2023_wasi} will include the support for asynchronous operations, which will also introduce new types such as \texttt{future<T>} and \texttt{stream<T, E>}. Our approach of benchmarking the serverless platforms can be extended to benchmark the asynchronous operations in WebAssembly. The future work can extensively evaluate I/O intensive workloads with WASI.
    \item \textbf{Security}: The security aspects of WebAssembly in serverless environment is a very important area of research. We have seen that AWS Lambda is using Firecracker microVM to run the serverless functions. The microVMs are chosen in favor of containers because of the security benefits they provide. Therefore, future research is essential to thoroughly evaluate the security facets of WebAssembly within a serverless environment.
\end{itemize}
